Exploring Rural Shrink Smart Through Guided Discovery Dashboards

### 2022-08-01 SDSS Journal Paper/Paper Draft.Rmd
Edits to remove the color edits

\section{ Discussion}
How did the philosophy work out?  
\db{The design philosophy worked well in the ways to promote guided discovery learning for the user. The blocks that were created helped create a way for the user to scroll to the next section. This framework in the design allowed for the highlighted pieces of the town in generally. This design allowed the user to see the PCP after understanding how the dashboard allowed for their town to centered. The PCP in the lower part of the dashboard allowed for the user to see the plot and adjust to the fact that they could add more towns to the plot.}

What worked and didn't (overall)?
\db{The idea to use the guided discovery methodology created a lack of understanding in the plot for a novice user. The need to explain a visual means that the user is more likely to not use the plot effectively. Our hope was that the PCP would be a bit more "self-explanatory" than it was initially. The number of helpful hints clutters the UI and may be a distraction for a particular user. }

People didn't seem to like having their community facts to explore from - they already knew that information - so what should we replace it with? Collaborator suggestions included reporting capability and data stories (narratives from the communities), but neither one really inspires exploratory data analysis. What we need to do is provide enough interactivity to allow people to select interesting data sets and use that to compare

\db{Many small towns don't like to see that their information is being displayed publicly or used to be compared to another city. This could a huge barrier for our resarch and work. To make sure that this does not become a problem, we could make sure that when allowing the user to compare towns in the PCP, we make sure to de-idenitify the towns in the UI. For example, the town user selects similiar population size, the each town that is populated will have a collection of ten similiar towns using a label, such as "Town 1", "Town 2", etc. }
\db{Along with the de-idenitifying other towns within the UI, using a login system for each town will allow for the town to feel as if their data is being protected. This login system can then be used as a filter to setup the dashboard there forward. This will hopefully remove the lack of ability to be interactive. This will seem as if the dashbaord is developed for their needs particularly.}
\db{Once, the privacy issue is removed to understand more of the data, we should allow for the user to add/remove variables in to the plot as they see fit. Suggesting that the QoL variables are the ones that are added in the start, but allowing the user to remove subjects from the QoL sections will allow for the user to not only feel as if the UI is interactive, it would also create more exploration for the user.}

We also should provide much, much more user support for how to interact with and interpret the app and results. 
\db{Support for the dashboard should have a help email for towns to feel supported with interpreting the app and results. This helpful email will be connected to our research group. In which the help will not only have statisticians but also city planners that can help answer questions. As town users are looking to make changes for their towns, it could be a QoL variable that will change the status of the town, which can be addressed with the statisticians. We will be able to help them with making sure that the right variables are being used in the PCPs. }

\section{ Future Work }
Incorporating other dashboards which had slightly different focuses (99 vs 990 towns, general data display, etc.).
\db{In our research, our future work will incorporate other dashboards with general information that are slightly different focuses (99 vs 990 towns, general data display, etc.). In collaboration with DSPG at Iowa State University and the statistical modeling portions of our NSF group, we would like to transition our app to a multi-tab dashboard which allows for more specific task-oriented exploration and evaluation, rather than broad EDA and discovery.}

Discussion of the difficulty of working with researchers from other disciplines, not because of convention but because they don't know what is possible with EDA and are unfamiliar with all but goal-directed data analysis.

\db{EDA can be a difficult for a variety of groups of people, novice users and experienced researchers. While working with a variety of  disciplines because EDA can be hard to understand the use cases of exploration of data. Many experience researchers, see EDA as a way to look at data that were collected to only describe the behavior of variables that are added into models or research questions. However, EDA can be used in a data discovery, such as the work displayed in this article. Data Discovery can help with the quality of the data, in order to find new resources. EDA can help with potential casuality of town related issues. For example, if data are showing that people are no longer using public transportation, this could be due to the population shrinking but exploaring the data could suggest that more people are using ride share companies. This could as suggest that public transportation may not offer the needs for the people in the town. EDA will help identify those issues or even allow for the town leaders to look for answers.}

\section {Conclusions}
Specific takeaways - visualization research (broad) vs. dashboard design (very problem specific) is missing a critical set of research describing the specific and trying to make general suggestions as to how to proceed with the design process.

\db{In conclusion, using a less known statistical graphic, our ideas of using Guided Dscovery Learning was used to make sure that the users were informed of the information that was presented, it did not allow for the town leaders any more ease in understanding. As a result, we decided to look for easier ways to display data, such as boxplots and barplots. This direction will be limited but may be useful for the sake of ease on the town leaders.}

\db{Complex visualizations are difficult to assess in vivo: it is fairly easy to design an experiment to test a single graphic, but much harder to test the interactive dashboards which consist of systems of displays and user interactions. As a result, it is important to continuously assess which ``common knowledge'' design principles stemming from more controlled experiments hold up in real-world situations.}

\db{In the event that we want to continue to have the town leaders explore using a different statistical graphic, we would like to introduce the flower plot. This method may not represent that data in the way that could be misleading for understanding, we believe that the flower plot will allow for town leaders to engage in being curious as they explore more dimensions of data related to their Quality of Life (QoL) survey.}

\db{With the lack of research in dashboard design, it could be challenging to develop a dashbaord with a clear layout and clear intention. This lack of clarity outside of dashboard pitfalls could lend the dashboard to miss key components that are useful. This is a generalization not only for our research but along all types of dashboard design. We are encouraged to minimize an interface that will not overwhelm the user, but what types of components, which components should be used can be a struggle for our research.}

Helpful hints - showing a collection of very different apps to train collaborators on what is possible works well to empower collaborators to help brainstorm.
Test early, test often...

\db{When showing our apps to internal research members, it was difficult to collective incoporate what was wanted and what was needed to empower collaborators to help brainstorm. Once, a design was presented, there were polarizing responses from no feedback to a dramatic change of the app layout. At times, collaborators could feel as if there suggestions were not being heard or dismissed. Along with the intention of the new designs could have been lost in translation due to the lack of being incoporated earlier or more often. As a result both groups become fustrated.  }

>>>>

\section{ Discussion}
Our dashboard design philosophy worked primarily to promote a town-centric approach application with comparisons to other similar towns being secondary. This approach created a way for the user to see their town information at the top of the page, this framework in the design allowed for the highlighted pieces of the town in generally. Once, the town components at the top of the page allowed the user to see the PCP after understanding their own status. The PCP in the lower part of the dashboard allowed for the user to see the plot and adjust to the fact that they could add more towns to the plot.

While the idea to use the guided discovery methodology created a lack of understanding in the plot for a novice user on the PCP. The need to explain a visual means that the user is more likely to not use the plot effectively. Our hope was that the PCP would be a bit more "self-explanatory" than it was initially. The number of helpful hints clutters the UI and may be a distraction from the data and the usefulness of the PCPs.

Another hurdle and underlying tone from rural towns would be the fact that a larger town would like to change the framework of the uniqueness of the town. As a result, small towns don't like to see that their information being displayed publicly or compared to another city. This could a huge barrier for our research and work. To make sure that this does not become a problem, in the future work making sure that when allowing the user to compare towns in a visual graphic with a de-identified towns in the UI. For example, the town user selects similar population size, the each town that is populated will have a collection of ten similar towns using a label, such as "Town 1", "Town 2", etc. Along with the de-identifying other towns within the UI, using a login system for each town will allow for the town to feel as if their data is being protected. This login system can then be used as a filter to setup the dashboard there forward. This will hopefully remove the lack of ability to be interactive in the purpose original design but give towns the comfort in feeling that their town data is being secure.

Once, the privacy issue is removed to understand more of the data, we should allow for the user to add/remove variables in to the plot as they see fit. By suggesting that the QoL variables are the ones that are added in the start, but allowing the user to remove subjects from the QoL sections will allow for the user to not only feel as if the UI is interactive, it would also create more exploration for the user.

Support for the dashboard application should have a help email for towns to ensure results and interpretation are clear. In this proposed idea, email communication will be connected to our research group, in which the help will not only have statisticians but also city planners that can help answer questions. Our integrated system will be able to help towns with data driven decision-making less of an invasion of privacy but as a collaborative workflow that will help their town succeed.

\section{ Future Work }
In our research, our future work will incorporate other dashboards with general information that are slightly different focuses (99 vs 990 towns, general data display, etc.). In collaboration with DSPG at Iowa State University and the statistical modeling portions of our NSF group, we would like to transition our app to a multi-tab dashboard which allows for more specific task-oriented exploration and evaluation, rather than broad EDA and discovery.

In our statistical graphs, we used Parallel Coordinate Plots (PCPs) to allow the user to look at a variety of variables, while the Guided Discovery Learning was used to makes sure that the users were informed of the information that was presented, it did not allow for the town leaders any more ease in understanding. This direction will be limited but may be useful for the sake of ease on the town leaders. In the event that we want to continue to have the town leaders explore using a different statistical graphic, we would like to introduce a flower plot. This method may not represent that data in the way that could be misleading for understanding, we believe that the flower plot will allow for town leaders to engage in being curious as they explore more dimensions of data related to their Quality of Life (QoL) survey.

EDA can be a difficult for a variety of groups of people, novice users and experienced researchers. While working with a variety of  disciplines because EDA can be hard to understand the use cases of exploration of data. Many experience researchers, see EDA as a way to look at data that were collected to only describe the behavior of variables that are added into models or research questions. However, EDA can be used in a data discovery, such as the work displayed in this article. Data Discovery can help with the quality of the data, in order to find new resources. EDA can help with potential causality of town related issues. For example, if data are showing that people are no longer using public transportation, this could be due to the population shrinking but exploring the data could suggest that more people are using ride share companies. This could as suggest that public transportation may not offer the needs for the people in the town. EDA will help identify those issues or even allow for the town leaders to look for answers.

\section {Conclusions}

In conclusion, using a less known statistical graphic, our ideas of using Guided Discovery Learning was used to make sure that the users were informed of the information that was presented, it did not allow for the town leaders any more ease in understanding. As a result, we decided to look for easier ways to display data, such as boxplots and barplots. This direction will be limited but may be useful for the sake of ease on the town leaders.

Complex visualizations are difficult to assess in vivo: it is fairly easy to design an experiment to test a single graphic, but much harder to test the interactive dashboards which consist of systems of displays and user interactions. As a result, it is important to continuously assess which ``common knowledge'' design principles stemming from more controlled experiments hold up in real-world situations.

In the event that we want to continue to have the town leaders explore using a different statistical graphic, we would like to introduce the flower plot. This method may not represent that data in the way that could be misleading for understanding, we believe that the flower plot will allow for town leaders to engage in being curious as they explore more dimensions of data related to their Quality of Life (QoL) survey.

With the lack of research in dashboard design, it could be challenging to develop a dashboard with a clear layout and clear intention. This lack of clarity outside of dashboard pitfalls could lend the dashboard to miss key components that are useful. This is a generalization not only for our research but along all types of dashboard design. We are encouraged to minimize an interface that will not overwhelm the user, but what types of components, which components should be used can be a struggle for our research.

When showing our apps to internal research members, it was difficult to collective incorporate what was wanted and what was needed to empower collaborators to help brainstorm. Once, a design was presented, there were polarizing responses from no feedback to a dramatic change of the app layout. At times, collaborators could feel as if there suggestions were not being heard or dismissed. Along with the intention of the new designs could have been lost in translation due to the lack of being incorporated earlier or more often. As a result both groups become frustrated.

### 2022-08-01 SDSS Journal Paper/Paper-Draft.tex
Edits to remove the color edits

How did the philosophy work out?\\
{\textcolor{blue}{The design philosophy worked well in the ways to promote guided discovery learning for the user. The blocks that were created helped create a way for the user to scroll to the next section. This framework in the design allowed for the highlighted pieces of the town in generally. This design allowed the user to see the PCP after understanding how the dashboard allowed for their town to centered. The PCP in the lower part of the dashboard allowed for the user to see the plot and adjust to the fact that they could add more towns to the plot.}}

What worked and didn't (overall)?
{\textcolor{blue}{The idea to use the guided discovery methodology created a lack of understanding in the plot for a novice user. The need to explain a visual means that the user is more likely to not use the plot effectively. Our hope was that the PCP would be a bit more "self-explanatory" than it was initially. The number of helpful hints clutters the UI and may be a distraction for a particular user. }}

People didn't seem to like having their community facts to explore from - they already knew that information - so what should we replace it with? Collaborator suggestions included reporting capability and data stories (narratives from the communities), but neither one really inspires exploratory data analysis. What we need to do is provide enough interactivity to allow people to select interesting data sets and use that to compare

{\textcolor{blue}{Many small towns don't like to see that their information is being displayed publicly or used to be compared to another city. This could a huge barrier for our resarch and work. To make sure that this does not become a problem, we could make sure that when allowing the user to compare towns in the PCP, we make sure to de-idenitify the towns in the UI. For example, the town user selects similiar population size, the each town that is populated will have a collection of ten similiar towns using a label, such as "Town 1", "Town 2", etc. }}
{\textcolor{blue}{Along with the de-idenitifying other towns within the UI, using a login system for each town will allow for the town to feel as if their data is being protected. This login system can then be used as a filter to setup the dashboard there forward. This will hopefully remove the lack of ability to be interactive. This will seem as if the dashbaord is developed for their needs particularly.}}
{\textcolor{blue}{Once, the privacy issue is removed to understand more of the data, we should allow for the user to add/remove variables in to the plot as they see fit. Suggesting that the QoL variables are the ones that are added in the start, but allowing the user to remove subjects from the QoL sections will allow for the user to not only feel as if the UI is interactive, it would also create more exploration for the user.}}

We also should provide much, much more user support for how to interact with and interpret the app and results.
{\textcolor{blue}{Support for the dashboard should have a help email for towns to feel supported with interpreting the app and results. This helpful email will be connected to our research group. In which the help will not only have statisticians but also city planners that can help answer questions. As town users are looking to make changes for their towns, it could be a QoL variable that will change the status of the town, which can be addressed with the statisticians. We will be able to help them with making sure that the right variables are being used in the PCPs. }}

@@

Incorporating other dashboards which had slightly different focuses (99 vs 990 towns, general data display, etc.).
{\textcolor{blue}{In our research, our future work will incorporate other dashboards with general information that are slightly different focuses (99 vs 990 towns, general data display, etc.). In collaboration with DSPG at Iowa State University and the statistical modeling portions of our NSF group, we would like to transition our app to a multi-tab dashboard which allows for more specific task-oriented exploration and evaluation, rather than broad EDA and discovery.}}

Discussion of the difficulty of working with researchers from other disciplines, not because of convention but because they don't know what is possible with EDA and are unfamiliar with all but goal-directed data analysis.

{\textcolor{blue}{EDA can be a difficult for a variety of groups of people, novice users and experienced researchers. While working with a variety of  disciplines because EDA can be hard to understand the use cases of exploration of data. Many experience researchers, see EDA as a way to look at data that were collected to only describe the behavior of variables that are added into models or research questions. However, EDA can be used in a data discovery, such as the work displayed in this article. Data Discovery can help with the quality of the data, in order to find new resources. EDA can help with potential casuality of town related issues. For example, if data are showing that people are no longer using public transportation, this could be due to the population shrinking but exploaring the data could suggest that more people are using ride share companies. This could as suggest that public transportation may not offer the needs for the people in the town. EDA will help identify those issues or even allow for the town leaders to look for answers.}}

@@

Specific takeaways - visualization research (broad) vs.~dashboard design (very problem specific) is missing a critical set of research describing the specific and trying to make general suggestions as to how to proceed with the design process.

{\textcolor{blue}{In conclusion, using a less known statistical graphic, our ideas of using Guided Dscovery Learning was used to make sure that the users were informed of the information that was presented, it did not allow for the town leaders any more ease in understanding. As a result, we decided to look for easier ways to display data, such as boxplots and barplots. This direction will be limited but may be useful for the sake of ease on the town leaders.}}

{\textcolor{blue}{Complex visualizations are difficult to assess in vivo: it is fairly easy to design an experiment to test a single graphic, but much harder to test the interactive dashboards which consist of systems of displays and user interactions. As a result, it is important to continuously assess which ``common knowledge'' design principles stemming from more controlled experiments hold up in real-world situations.}}

{\textcolor{blue}{In the event that we want to continue to have the town leaders explore using a different statistical graphic, we would like to introduce the flower plot. This method may not represent that data in the way that could be misleading for understanding, we believe that the flower plot will allow for town leaders to engage in being curious as they explore more dimensions of data related to their Quality of Life (QoL) survey.}}

{\textcolor{blue}{With the lack of research in dashboard design, it could be challenging to develop a dashbaord with a clear layout and clear intention. This lack of clarity outside of dashboard pitfalls could lend the dashboard to miss key components that are useful. This is a generalization not only for our research but along all types of dashboard design. We are encouraged to minimize an interface that will not overwhelm the user, but what types of components, which components should be used can be a struggle for our research.}}

Helpful hints - showing a collection of very different apps to train collaborators on what is possible works well to empower collaborators to help brainstorm.
Test early, test often\ldots{}

{\textcolor{blue}{When showing our apps to internal research members, it was difficult to collective incoporate what was wanted and what was needed to empower collaborators to help brainstorm. Once, a design was presented, there were polarizing responses from no feedback to a dramatic change of the app layout. At times, collaborators could feel as if there suggestions were not being heard or dismissed. Along with the intention of the new designs could have been lost in translation due to the lack of being incoporated earlier or more often. As a result both groups become fustrated.  }}

>>>>

Our dashboard design philosophy worked primarily to promote a town-centric approach application with comparisons to other similar towns being secondary. This approach created a way for the user to see their town information at the top of the page, this framework in the design allowed for the highlighted pieces of the town in generally. Once, the town components at the top of the page allowed the user to see the PCP after understanding their own status. The PCP in the lower part of the dashboard allowed for the user to see the plot and adjust to the fact that they could add more towns to the plot.

While the idea to use the guided discovery methodology created a lack of understanding in the plot for a novice user on the PCP. The need to explain a visual means that the user is more likely to not use the plot effectively. Our hope was that the PCP would be a bit more ``self-explanatory'' than it was initially. The number of helpful hints clutters the UI and may be a distraction from the data and the usefulness of the PCPs.

Another hurdle and underlying tone from rural towns would be the fact that a larger town would like to change the framework of the uniqueness of the town. As a result, small towns don't like to see that their information being displayed publicly or compared to another city. This could a huge barrier for our research and work. To make sure that this does not become a problem, in the future work making sure that when allowing the user to compare towns in a visual graphic with a de-identified towns in the UI. For example, the town user selects similar population size, the each town that is populated will have a collection of ten similar towns using a label, such as ``Town 1'', ``Town 2'', etc. Along with the de-identifying other towns within the UI, using a login system for each town will allow for the town to feel as if their data is being protected. This login system can then be used as a filter to setup the dashboard there forward. This will hopefully remove the lack of ability to be interactive in the purpose original design but give towns the comfort in feeling that their town data is being secure.

Once, the privacy issue is removed to understand more of the data, we should allow for the user to add/remove variables in to the plot as they see fit. By suggesting that the QoL variables are the ones that are added in the start, but allowing the user to remove subjects from the QoL sections will allow for the user to not only feel as if the UI is interactive, it would also create more exploration for the user.

Support for the dashboard application should have a help email for towns to ensure results and interpretation are clear. In this proposed idea, email communication will be connected to our research group, in which the help will not only have statisticians but also city planners that can help answer questions. Our integrated system will be able to help towns with data driven decision-making less of an invasion of privacy but as a collaborative workflow that will help their town succeed.

@@

In our research, our future work will incorporate other dashboards with general information that are slightly different focuses (99 vs 990 towns, general data display, etc.). In collaboration with DSPG at Iowa State University and the statistical modeling portions of our NSF group, we would like to transition our app to a multi-tab dashboard which allows for more specific task-oriented exploration and evaluation, rather than broad EDA and discovery.

In our statistical graphs, we used Parallel Coordinate Plots (PCPs) to allow the user to look at a variety of variables, while the Guided Discovery Learning was used to makes sure that the users were informed of the information that was presented, it did not allow for the town leaders any more ease in understanding. This direction will be limited but may be useful for the sake of ease on the town leaders. In the event that we want to continue to have the town leaders explore using a different statistical graphic, we would like to introduce a flower plot. This method may not represent that data in the way that could be misleading for understanding, we believe that the flower plot will allow for town leaders to engage in being curious as they explore more dimensions of data related to their Quality of Life (QoL) survey.

EDA can be a difficult for a variety of groups of people, novice users and experienced researchers. While working with a variety of disciplines because EDA can be hard to understand the use cases of exploration of data. Many experience researchers, see EDA as a way to look at data that were collected to only describe the behavior of variables that are added into models or research questions. However, EDA can be used in a data discovery, such as the work displayed in this article. Data Discovery can help with the quality of the data, in order to find new resources. EDA can help with potential causality of town related issues. For example, if data are showing that people are no longer using public transportation, this could be due to the population shrinking but exploring the data could suggest that more people are using ride share companies. This could as suggest that public transportation may not offer the needs for the people in the town. EDA will help identify those issues or even allow for the town leaders to look for answers.

@@

In conclusion, using a less known statistical graphic, our ideas of using Guided Discovery Learning was used to make sure that the users were informed of the information that was presented, it did not allow for the town leaders any more ease in understanding. As a result, we decided to look for easier ways to display data, such as boxplots and barplots. This direction will be limited but may be useful for the sake of ease on the town leaders.

Complex visualizations are difficult to assess in vivo: it is fairly easy to design an experiment to test a single graphic, but much harder to test the interactive dashboards which consist of systems of displays and user interactions. As a result, it is important to continuously assess which ``common knowledge'\,' design principles stemming from more controlled experiments hold up in real-world situations.

In the event that we want to continue to have the town leaders explore using a different statistical graphic, we would like to introduce the flower plot. This method may not represent that data in the way that could be misleading for understanding, we believe that the flower plot will allow for town leaders to engage in being curious as they explore more dimensions of data related to their Quality of Life (QoL) survey.

With the lack of research in dashboard design, it could be challenging to develop a dashboard with a clear layout and clear intention. This lack of clarity outside of dashboard pitfalls could lend the dashboard to miss key components that are useful. This is a generalization not only for our research but along all types of dashboard design. We are encouraged to minimize an interface that will not overwhelm the user, but what types of components, which components should be used can be a struggle for our research.

When showing our apps to internal research members, it was difficult to collective incorporate what was wanted and what was needed to empower collaborators to help brainstorm. Once, a design was presented, there were polarizing responses from no feedback to a dramatic change of the app layout. At times, collaborators could feel as if there suggestions were not being heard or dismissed. Along with the intention of the new designs could have been lost in translation due to the lack of being incorporated earlier or more often. As a result both groups become frustrated.
